### 2022-11-05 NSF2022/Description/backgroundv2.tex
working on background

\begin{figure}
  \begin{center}
    \includegraphics[width=0.45\textwidth]{figures/SpecificAim1/Domain.jpg}
% \includegraphics[width=0.55\textwidth]{figures/SpecificAim1/Domain.jpg}    
 \includegraphics[width=0.45\textwidth]{figures/SpecificAim1/3Particles.jpg}
  \end{center}
  \caption{\label{fig:flow_map}
   Left: In the HAP formulation,
   a system of exterior problems 
   defines the hydronamic and hydrophobic interactions.
  Right:
     Particles translate and rotate to lower the free energy $F[u]$
   subject to \eqref{eq:SL}, $f(u) = u^2$, and reach a minimal configuration in
   panel C. The colormap is blue for $u = 0$ and red for $u = 1$.
   The arrows are the velocity field $\mathbf{u}$.}
\end{figure}

\subsection{Problem formulation}
Recently, we have developed a new mathematical model of self-organization.
The model takes ideas from the phase field method 
and leverages boundary integral equations to speed up computations.
To formulate the problem, consider a collection of $N_b$-many rigid bodies
$U_i$ immersed in an incompressible, zero-Reynolds number fluid
$\Omega_t \mathbb{R}^n \setminus \cup_{i=1}^{N_b} U_i$.
The subscript $t$ reinforces that $\Omega_t$ is time-varying. 
We use the grain shape and boundary conditions to capture the local 
properties of the suspended media.
Unlike particles, the bodies, which we will call grains (``using `grains' as an easy find-replace for
when we settle on vocabulary), have finite size.

The first key idea is to define the grain interactions using a boundary
value problem
\begin{alignat}{3}
  -\mu\Delta \uu + \nabla p &= \mathbf{0}, \:
  \nabla \cdot \uu = 0, \quad &&\xx \in \Omega, \label {eqn:stokes} \\
  -\rho^2 \Delta \eta + f'(\eta) &= 0, &&\xx \in \Omega,
  \label{eqn:phase}
\end{alignat}
with boundary conditions
\begin{alignat}{3}
  \frac{d\xx}{dt} &= \uu, &&\xx \in \bd\Omega \\
  \uu &= \vv_i + \omega_i \times (\xx - \aa_i), \quad &&\xx \in
  \Gamma_i,\, i=1,\ldots,N_b, \\
  \eta &= h_i(\xx,t),  &&\xx \in \Gamma_i,\, i=1,\ldots,N_b,
\end{alignat}
and
\begin{equation}
\int_{\Gamma_i} \mu(\nabla \uu + \nabla \uu^T + pI)\nu \,\dif s = 
\int_{\Gamma_i}
\underbrace{(-\rho^{-1} f'(\eta)
  + \rho(\tfrac{1}{2}|\nabla \eta|^2 - \nabla \eta \nabla \eta^T)}_{T_i} \nu \, \dif s
\end{equation}
This is a semi-linear system for evolution of the grains.  At the top,
the Stokes equations describe the velocity $\uu$ and pressure $p$ of the
aqueous phase. In third equation describes the phase transitions of the
scalar order parameter $\eta$ with a decay length $\rho.$
Mathematically, this order parameter is responsible for the attraction
and adhesion between grains that want to align boundaries of like phase.  
Physically, water molecules are polar and $\eta$ describes the orientational
order of water e.g. $\eta = 1$ when water molecules share alignment
and $\eta = 0$ when the water orientations are disordered.
The well-potential $f$ accounts for possibly distinct equilibrium phases,
as occurs with 3-bond and 4-bond hydrogen bonding.

The next three equations are the moving domain boundary conditions.
The first two specify a non-slip boundary condition for a rigid body
with translation velocity $\vv_i$ and angular velocity $\omega_i$.
In the third boundary condition, the function $h_i$ is a material label
speficying the water phase at the water-grain interface.  This material
label is transported by the fluid and therefore satisfies $h_i(\xx,t)
= h^0_i(\xx(\XX,t),t)$ where $h_i^0$ and $\XX$ are reference labels, coordinates
given by the initial conditions, and $\xx(\XX,t)$ is the flow map for the
rigid bodies.  Finally, the last boundary condition closes the system
and says that the hydrodynamic stress is balanced by the water-order stress.

Intriguingly, the formulation ()-() captures exactly the right behavior
expected from amphiphilic self-assembly.  Three bodies with amphiphilic
label spontaneously translate and change orientations to sequester the
disordered fluid phase.  In our simulations (see section ()), we include
a finite length repulsion to avoid particle collisions.
Moreover, a simple change of boundary conditions from amphiphilic,
to biased hydrophobic, to polar results in distinct equilibrium morphologies.

\begin{figure}[h]
  \begin{center}
    \begin{tabular}{m{0.1cm}m{1.8in}m{1.8in}m{1.8in}}            
      (a)
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100B1.jpg}
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100B2.jpg}
     &\includegraphics[width=1.8in]{figures/SpecificAim1/N100B3.jpg}    \\
     (b)
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100C1.jpg}
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100C2.jpg}
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100C3.jpg}    \\
    (c)
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100A1.jpg}
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100A2.jpg}
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100A3.jpg}    \\
      &\multicolumn{1}{c}{$t = 0$}
      &\multicolumn{1}{c}{intermediate time}
      &\multicolumn{1}{c}{long time}
  \end{tabular}
  \end{center}
  \vspace{-20pt}
  \caption{\footnotesize
    \label{fig:self-assembly}
    Rows (a)-(c) have the same initial configuration
    but have different boundary conditions $g.$ 
    In row (a), self-assembly with an amphiphilic boundary
    condition \eqref{eq:bcs}(a) results in a bilayer pattern shielding
    the hydrophobic core (yellow, $u > 0$) from the aqueous phase (blue, $u = 0$).
    Row (b) shows particles with a hydrophobic intensity that is
    greater on one side of the particle (magenta) than the other (yellow).
    In row (c), green is for $u < 0$, yellow is for $u > 0$, and blue is for $u = 0$.
    Particles orient like sides to form a checkerboard pattern.\\
  }
\end{figure}

The system ()-() has an elegant mathematical structure.
In terms of locally in time well-posedness, given disjoint grains with
sufficiently smooth boundaries
and data, the system ()-() always has a solution and that these solutions
have a locally Lipschitz dependence on the particle centers and orientations. 
Solutions of ()-() obey the energy law
\begin{equation}
  \frac{d}{dt}F
  = - \int_{\Omega_t} \tfrac{1}{2}\mu|\nabla \uu + \nabla \uu^T|^2 \,dx,
  \quad
  F = \int_{\Omega_t}
  \rho |\nabla \eta|^2 + \rho^{-1} f(\eta) \,dx.
\end{equation}
Mathematically, this energy law states that the free energy $F$ of the water order
phase transitions, which depends on the grain configurations, is lost due to
hydrodynamic dissipation.  It is evident that minimizers of the functional $F$
are unique and satisfy the Euler-Lagrange equation ().
There is a subtlety in the energy law, though.  Namely, the
time derivative of the free energy $F$ also involves the moving
boundary, accounted for by the Rayleigh Transport Theorem, and the
fact that the boundary data for $\eta$ are transported by the rigid body motion.
In fact, the water order surface stress $T_i$ is the variational derivative that
shows up in this calculation.

@@

To summarize, the formulation ()-() shares a lot of the features in phase field modeling,
with the big exception that the free energy depends on the grain configurations,
rather than phase field
function.
To make an explicit connection though, we can consider a phase field model 
where one ``fluid'' phase instantaneously reaches minimality while the other ``solid''
phase moves to further decrease free energy. 
The model is also closely related to
a recent and strongly developing direction studying 
colloids immersed in nematic liquid crystals.
As is the case for nematics, it is natural to understand what is the effect of the flow in
such a context, but a first step is to obtain some solid well-posedness results.
The proposal discusses some mathematical generalizations in section ().

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary Integral formulation}
Boundary integral equations (BIEs) are a reformulation of elliptic PDEs.
For example, solutions of the Stokes equations~\eqref{eqn:stokes} can be
written as the convolution 
\begin{align*}
  \uu(\xx) = \int_{\partial\Omega} G(\xx - \yy) \ssigma(\yy) \, ds_\yy,
\end{align*}
where $G(\xx - \yy)$ is a carefully chosen derivative of the PDE's
fundamental solution and $\ssigma: \partial \Omega \rightarrow \RR^2$ is
an unknown density function. The BIE reformulation has several immediate
advantages including a dimension reduction since the unknown density
function is defined only on $\partial \Omega$, and the far field
boundary condition of the PDE is automatically satisfied. The density
function $\ssigma$ is chosen to satisfy the boundary condition, and this
results in a BIE 
\begin{align*}
  \gg(\xx) = \lambda \ssigma(\xx) + 
    \int_{\partial\Omega} G(\xx - \yy) \ssigma(\yy) \, ds_\yy,
    \quad \xx \in \bd\Omega,
\end{align*}
where $\gg$ is the Dirichlet boundary condition, and $\lambda$ is a
constant that depends on the choice of $G$. Note that if the
well-potential is $f(u) = \tfrac{1}{2}u^2$, then
equation~\eqref{eqn:phase} becomes the so-called linear screened Laplace
equation and can be written as a BIE with an unknown scalar-valued
density function.

@@

%The boundary integral equation (BIE) method is the ideal numerical
%method for the present study in terms of computational cost and
%accuracy. Boundary integral equations require linear PDE and so we take
%$f(\eta) = \tfrac{1}{2}\eta^2$, yielding in () the so called screened
%Laplace equation $-\rho^2 \Delta \eta + \eta$.  This is a Helmholtz
%equation with purely imaginary wave number, enabling us to represent the
%solution in terms of layer potentials:
%\begin{equation}
%  \eta(\xx) = \int_{\partial \Omega} \frac{\partial G}{\partial \nu}
%  \left(\frac{\xx-\yy}{\rho}\right) \sigma(\yy) \,ds_{\yy}
%\end{equation}
%where $\frac{\partial G}{\partial \nu}$ is the normal derivative of the
%fundamental solution of the screened Laplace equation
%(a Bessel function in two space dimensions and
%an exponential times the standard fundamental solution of the Laplace equation
%in three space dimensions).
%This representation allows us to transfer the boundary value problem ()-()
%to an equation only involving in terms of domain boundary.  
%
%The numerical advantages of doing so are as follows.  If $h = N^{-1}$ is
%the characeristic grid spacing of the grain boundaries, then a finite
%element or finite difference discretization of ()-() requires on the
%order of $N_b N^n$ grid points for dense suspensions.  In contrast BIE
%require onely With finite elements, mesh regeneration would be needed at
%each time step since the domain is moving.  In the boundary integral
%approach, collocation points can be selected for the initial grain
%configuration and then used throughout the time-course.  As a downside,
%integral equations require specialized quadrature routines to resolve
%self interactions and nearly touching grain boundaries.  Finally, the
%far field boundary conditions which are important for imposed flow
%problems automatically hold, whereas with finite element and finite
%difference methods artificial boundary conditions must be imposed.
\begin{wrapfigure}[10]{r}{2in}
  \todo[inline]{A figure comparing FEM versus BIE meshes}
  \caption{\label{fig:fem_vs_bie} A finite element method requires
  meshing the entire domain (typically into triangles), while a BIE
  only requires meshing the boundary of the domain.}
\end{wrapfigure}

>>>>

\begin{figure}
  \begin{center}
    \includegraphics[width=0.45\textwidth]{figures/SpecificAim1/Domain.jpg}
% \includegraphics[width=0.55\textwidth]{figures/SpecificAim1/Domain.jpg}    
 \includegraphics[width=0.45\textwidth]{figures/SpecificAim1/3Particles.jpg}
  \end{center}
  \caption{\label{fig:flow_map} \footnotesize {\em Left:} In the HAP
  formulation, a system of exterior problems defines the hydrodynamic
  and hydrophobic interactions. {\em Right:} Particles translate and
  rotate to lower the free energy $F[u]$ subject to \eqref{eq:SL}, $f(u)
  = u^2$, and reach a minimal configuration in panel C. The colormap is
  blue for $u = 0$ and red for $u = 1$. The arrows are the velocity
  field $\mathbf{u}$.}
\end{figure}

\subsection{Problem formulation}
Recently, we have developed a new mathematical model of
self-organization~\cite{FuQuRyYo22,ru-ryh-qua-you2022,Fu2018_SIAM}. The
model takes ideas from the phase field method and leverages boundary
integral equations to speed up computations. To formulate the problem,
consider a collection of $N_b$-many rigid bodies $U_i$, with boundary
$\Gamma_i$, immersed in an incompressible, zero-Reynolds number fluid
$\Omega_t = \RR^n \setminus \cup_{i=1}^{N_b} U_i$. The subscript $t$
reinforces that $\Omega_t$ is time-varying. We use the grain shape and
boundary conditions to capture the local properties of the suspended
media. Unlike particles, the bodies, which we will call grains, have
finite size.  \todo[inline]{using `grains' as an easy find-replace for
when we settle on vocabulary}

The first key idea is to define the grain interactions using a boundary
value problem
\begin{alignat}{3}
  -\mu\Delta \uu + \nabla p &= \mathbf{0}, \:
  \nabla \cdot \uu = 0, \quad &&\xx \in \Omega_t, \label {eqn:stokes} \\
  -\rho^2 \Delta \eta + f'(\eta) &= 0, &&\xx \in \Omega_t,
  \label{eqn:phase}
\end{alignat}
with boundary conditions
\begin{alignat}{3}
  \frac{d\xx}{dt} &= \uu, &&\xx \in \bd\Omega_t \label{eqn:noslip} \\
  \uu &= \vv_i + \omega_i \times (\xx - \aa_i), \quad &&\xx \in
  \Gamma_i,\, i=1,\ldots,N_b, \label{eqn:rbm} \\
  \eta &= h_i(\xx,t),  &&\xx \in \Gamma_i,\, i=1,\ldots,N_b,
  \label{eqn:material}
\end{alignat}
and
\begin{align}
\int_{\Gamma_i} \mu(\nabla \uu + \nabla \uu^T + pI)\nu \,\dif s = 
\int_{\Gamma_i}
\underbrace{\left(-\rho^{-1} f'(\eta)
  + \rho\left(\tfrac{1}{2}|\nabla \eta|^2 - \nabla \eta \nabla
  \eta^T\right)\right)}_{T_i}
  \nnu \, \dif s.
  \label{eqn:stressbalance}
\end{align}
This is a semi-linear system for evolution of the grains.
Equation~\eqref{eqn:stokes} is the Stokes equations that describes the
velocity $\uu$ and pressure $p$ of the viscous aqueous phase.
Equation~\eqref{eqn:phase} describes the phase transitions of the scalar
order parameter $\eta$ with a decay length $\rho$. Mathematically, this
order parameter is responsible for the attraction and adhesion between
grains that want to align boundaries of like phase. Physically, water
molecules are polar and $\eta$ describes the orientational order of
water e.g.~$\eta = 1$ when water molecules share alignment and $\eta =
0$ when the water orientations are disordered. The well-potential $f$
accounts for possibly distinct equilibrium phases, as occurs with 3-bond
and 4-bond hydrogen bonding.

The next three equations are the moving domain boundary conditions.
Equations~\eqref{eqn:noslip} and~\eqref{eqn:rbm} specify a non-slip
boundary condition for a rigid body with translation velocity $\vv_i$
and angular velocity $\omega_i$ about the point $\aa_i \in \Omega_i$. In
equation~\eqref{eqn:material}, the function $h_i$ is a material label
specifying the water phase at the water-grain interface. This material
label is transported by the fluid and therefore satisfies $h_i(\xx,t) =
h^0_i(\xx(\XX,t),t)$ where $h_i^0$ and $\XX$ are reference labels,
coordinates given by the initial conditions, and $\xx(\XX,t)$ is the
flow map for the rigid bodies. Finally,
equation~\eqref{eqn:stressbalance} closes the system by requiring that
the hydrodynamic stress is balanced by the water-order stress.

Intriguingly, the
formulation~\eqref{eqn:stokes}--\eqref{eqn:stressbalance} captures
exactly the right behavior expected from amphiphilic self-assembly.
Three bodies with amphiphilic label spontaneously translate and change
orientations to sequester the disordered fluid phase. In our simulations
(see section ()), we include a finite length repulsion to avoid particle
collisions. Moreover, a simple change of boundary conditions from
amphiphilic, to biased hydrophobic, to polar results in distinct
equilibrium morphologies (Figure~\ref{fig:self-assembly}).

\begin{figure}[h]
  \begin{center}
    \begin{tabular}{m{0.1cm}m{1.8in}m{1.8in}m{1.8in}}            
      (a)
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100B1.jpg}
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100B2.jpg}
     &\includegraphics[width=1.8in]{figures/SpecificAim1/N100B3.jpg}    \\
     (b)
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100C1.jpg}
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100C2.jpg}
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100C3.jpg}    \\
    (c)
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100A1.jpg}
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100A2.jpg}
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100A3.jpg}    \\
      &\multicolumn{1}{c}{$t = 0$}
      &\multicolumn{1}{c}{intermediate time}
      &\multicolumn{1}{c}{long time}
  \end{tabular}
  \end{center}
  \vspace{-20pt}
  \caption{\footnotesize \label{fig:self-assembly} Rows (a)-(c) have the
  same initial configuration but different boundary conditions $g$. In
  row (a), self-assembly with an amphiphilic boundary condition results
  in a bilayer pattern shielding the hydrophobic core (yellow, $u > 0$)
  from the aqueous phase (blue, $u = 0$). Row (b) shows particles with a
  hydrophobic intensity that is greater on one side of the particle
  (magenta) than the other (yellow). In row (c), green is for $u < 0$,
  yellow is for $u > 0$, and blue is for $u = 0$. Particles orient like
  sides to form a checkerboard pattern.}
\end{figure}

The system~\eqref{eqn:stokes}--\eqref{eqn:stressbalance} has an elegant
mathematical structure. In terms of locally in time well-posedness,
given disjoint grains with sufficiently smooth boundaries and data, the
system always has a solution and these solutions have a locally
Lipschitz dependence on the particle centers and orientations.
Solutions of~\eqref{eqn:stokes}--\eqref{eqn:stressbalance} obey the
energy law
\begin{equation}
  \frac{d}{dt}F
  = - \int_{\Omega_t} \tfrac{1}{2}\mu|\nabla \uu + \nabla \uu^T|^2 \,d\xx,
  \quad
  F = \int_{\Omega_t}
  \rho |\nabla \eta|^2 + \rho^{-1} f(\eta) \,d\xx.
\end{equation}
Mathematically, this energy law states that the free energy $F$ of the
water order phase transitions, which depends on the grain
configurations, is lost due to hydrodynamic dissipation. It is evident
that minimizers of the functional $F$ are unique and satisfy the
Euler-Lagrange equation (). There is a subtlety in the energy law,
though. Namely, the time derivative of the free energy $F$ also involves
the moving boundary, accounted for by the Rayleigh Transport Theorem,
and the fact that the boundary data for $\eta$ are transported by the
rigid body motion. In fact, the water order surface stress $T_i$ is the
variational derivative that shows up in this calculation.

@@

To summarize, the
formulation~\eqref{eqn:stokes}--\eqref{eqn:stressbalance} shares a lot
of the features in phase field modeling, with the big exception that the
free energy depends on the grain configurations, rather than phase field
function. To make an explicit connection though, we can consider a phase
field model where one ``fluid'' phase instantaneously reaches minimality
while the other ``solid'' phase moves to further decrease free energy.
The model is also closely related to a recent and strongly developing
direction studying colloids immersed in nematic liquid crystals. As is
the case for nematics, it is natural to understand what is the effect of
the flow in such a context, but a first step is to obtain some solid
well-posedness results. The proposal discusses some mathematical
generalizations in section ().

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary Integral formulation}
Boundary integral equations (BIEs) are a reformulation of elliptic PDEs.
For example, solutions of the Stokes equations~\eqref{eqn:stokes} can be
written as the convolution 
\begin{align*}
  \uu(\xx) = \int_{\partial\Omega_t} G(\xx - \yy) \ssigma(\yy) \, ds_\yy,
\end{align*}
where $G(\xx - \yy)$ is a carefully chosen derivative of the PDE's
fundamental solution and $\ssigma: \partial \Omega_t \rightarrow \RR^2$
is an unknown density function. The BIE reformulation has several
immediate advantages including a dimension reduction since the unknown
density function is defined only on $\partial \Omega_t$, and the far
field boundary condition of the PDE is automatically satisfied. The
density function $\ssigma$ is chosen to satisfy the boundary condition,
and this results in a BIE 
\begin{align*}
  \gg(\xx) = \lambda \ssigma(\xx) + 
    \int_{\partial\Omega_t} G(\xx - \yy) \ssigma(\yy) \, ds_\yy,
    \quad \xx \in \partial\Omega_t,
\end{align*}
where $\gg$ is the Dirichlet boundary condition, and $\lambda$ is a
constant that depends on the choice of $G$. Note that if the
well-potential is $f(u) = \tfrac{1}{2}u^2$, then
equation~\eqref{eqn:phase} becomes the so-called linear screened Laplace
equation and can be written as a BIE with an unknown scalar-valued
density function.

@@

%The boundary integral equation (BIE) method is the ideal numerical
%method for the present study in terms of computational cost and
%accuracy. Boundary integral equations require linear PDE and so we take
%$f(\eta) = \tfrac{1}{2}\eta^2$, yielding in () the so called screened
%Laplace equation $-\rho^2 \Delta \eta + \eta$.  This is a Helmholtz
%equation with purely imaginary wave number, enabling us to represent the
%solution in terms of layer potentials:
%\begin{equation}
%  \eta(\xx) = \int_{\partial \Omega} \frac{\partial G}{\partial \nu}
%  \left(\frac{\xx-\yy}{\rho}\right) \sigma(\yy) \,ds_{\yy}
%\end{equation}
%where $\frac{\partial G}{\partial \nu}$ is the normal derivative of the
%fundamental solution of the screened Laplace equation
%(a Bessel function in two space dimensions and
%an exponential times the standard fundamental solution of the Laplace equation
%in three space dimensions).
%This representation allows us to transfer the boundary value problem ()-()
%to an equation only involving in terms of domain boundary.  
%
%The numerical advantages of doing so are as follows.  If $h = N^{-1}$ is
%the characeristic grid spacing of the grain boundaries, then a finite
%element or finite difference discretization of ()-() requires on the
%order of $N_b N^n$ grid points for dense suspensions.  In contrast BIE
%require onely With finite elements, mesh regeneration would be needed at
%each time step since the domain is moving.  In the boundary integral
%approach, collocation points can be selected for the initial grain
%configuration and then used throughout the time-course.  As a downside,
%integral equations require specialized quadrature routines to resolve
%self interactions and nearly touching grain boundaries.  Finally, the
%far field boundary conditions which are important for imposed flow
%problems automatically hold, whereas with finite element and finite
%difference methods artificial boundary conditions must be imposed.
\begin{wrapfigure}[10]{r}{2in}
  \todo[inline]{A figure comparing FEM versus BIE meshes}
  \caption{\label{fig:fem_vs_bie} \footnotesize A finite element method
  requires meshing the entire domain (typically into triangles), while a
  BIE only requires meshing the boundary of the domain.}
\end{wrapfigure}

### 2022-11-05 NSF2022/Description/specificaim2.tex
pushed organizational changes
started to work on BIE section
working on background

\subsection{Specific Aim 2: Efficient, high-order methods for
large-scale simulations}
% ----------------------------------------------------------------------
Over the past decade, there has been an explosion of interest in
small-scale processes that utilize capillary forces and van der Waals
interactions to coordinate movement and bind microscopic components in
solvent~\cite{Pandey2011, Zhang2017, Siontorou2017}. Additionally,
hydrodynamic interactions are central to fabricate complex,
three-dimensional microstructures~\cite{Dasgupta2017, Leong2007,
Reynolds2019, Cho2010}. Hydrodynamic effects cannot be ignored since
they, like for the rates of biological functions like pore
dynamics~\cite{RYHAM20112929}, set the rate of dissipation. In Specific
Aim 2, we propose novel and efficient high-order numerical methods to
simulate dynamic assembly of particles under the HAP in a viscous
solvent. As described in~\eqref{eq:stokes} and illustrated in
Figure~\ref{fig:flow_map}, the solvent phase is modeled by the Stokes
equations for an incompressible, viscous fluid, and these equations are
coupled to the screened-Laplace equation~\eqref{eq:SL} through viscous
and hydrophobic stress balance.

The HAP model requires solving exterior problems for the order
parameter~\eqref{eq:SL} and the Stokes equations~\eqref{eq:stokes}. We
will develop methods based on combinations of volume and layer
potentials, and inexact Newton's method. Numerical methods to solve
these integral equations (IEs) have several advantages over their
PDE-based counterpart: by construction, they satisfy far-field and mass
conservation conditions; only $\partial\Omega$ needs to be discretized
for layer potentials; carefully chosen IEs result in a well-conditioned
linear system that is solved with a mesh-independent number of GMRES
iterations; and high-order or spectral accuracy is attainable with
appropriate quadrature methods. The PIs have demonstrated that IEs are a
powerful tool to simulate two-dimensional suspensions of amphiphilic
particles~\cite{Fu2018_SIAM, FuQuRyYo20}. Moreover, elements of these
formulations have been described in three dimensions~\cite{ying_2006,
manasthesis, rac-gre2016}. This proposal will extend the two-dimensional
methods to more general amphiphilic particle models that
satisfy~\eqref{eq:SL}, and all the methods can naturally be extended to
three dimensions. These methods will be motivated by models and analysis
developed in Specific Aims 1 and 3. \S\ref{subsec:AC} describes how the
non-linear Allen-Cahn equation~\eqref{eq:SL} will be iteratively solved
using an IE formulation. Then, \S\ref{subsec:NumericalIssues} describes
how we will combine preconditioning strategies with fast summation
methods to accelerate the time to solution, and
\S\ref{subsec:timeStepping} describes adaptive time stepping methods.

@@

Having demonstrated that the inexact Newton method converges in one
dimension, it will be coupled with IE formulations to 
solve~\eqref{eq:SL} in two and three dimensions. We first partition
$\Omega$ into $\Omega_0 = \{\xx \in \Omega \:|\: u(\xx) < (u_0 +
u_1)/2\}$ and $\Omega_1 = \{\xx \in \Omega \:|\: u(\xx) > (u_0 +
u_1)/2\}$, and we let $\Lambda = \{\xx \in \Omega \:|\: u(\xx) = (u_0 +
u_1)/2 \}$ be their common boundary, and we let $G_0$ be the fundamental
solution of $(-\rho^2 \Delta + a)$ and $G_1$ be the fundamental
solution of $(-\rho^2 \Delta + 1)$. Then, the solution
of~\eqref{eq:screenedPoisson} can be written as $u = u^P + u^H$ where
$u^P$ is a volume potential and $u^H$ is a layer potential. The volume
potential is
\begin{align*}
  u^{P}(\xx) = \int_{\Omega_i} G_i(\xx,\yy) J[u^N](\yy)\, d\yy, 
    \quad i = 1,2,
\end{align*}
but it is discontinuous across $\Lambda$ and does not satisfy the
prescribed hydrophobic boundary conditions on the particles $\Gamma_i$.
Assuming the particles are entirely contained in the hydrophobic region,
the boundary conditions will be enforced with the layer potentials
\begin{align*}
  u^{H}(\xx) = \begin{cases}
    \displaystyle\int_{\Lambda} G_0(\xx,\yy) \eta_0(\yy) \, ds_\yy + 
    \displaystyle\int_{\Gamma} \pderiv{}{\nn_\yy} G_0(\xx,\yy)
      \sigma(\yy) \,ds_\yy, \quad & \xx \in \Omega_0, \\
    \displaystyle\int_{\Lambda} G_1(\xx,\yy) \eta_1(\yy) \, ds_\yy,
     & \xx \in \Omega_1.
  \end{cases}
\end{align*}
The density functions $\sigma$, $\eta_0$, and $\eta_1$ are determined so
that $u = u^P + u^H$ satisfies the hydrophobic boundary condition on
$\Gamma = \cup_i \Gamma_i$, and that it is continuous and differentiable
across $\Lambda$. This approach of using IEs to solve elliptic PDEs with
piecewise constant coefficients has successfully been used in other
application such as scattering~\cite{hyu-bar2014, che-cho-cai2018}.

>>>>

\subsection{Specific Aim 2: Efficient, high-order methods for
large-scale simulations}
\label{sec:specificaim2}
\begin{itemize}
  \item Make clear that we know how to address NSI and fast summation
  \item Talk about collision as per talk with Bryce
  \item Wen's paper can be used to give numbers on how large or a
    problem could hypothetically be considered
  \item Emphasize that we've learned a lot from just $\mathcal(O)(100)$
    particles, but now it's time to do more
  \item Confinement and rheology
\end{itemize}

% ----------------------------------------------------------------------
Over the past decade, there has been an explosion of interest in
small-scale processes that utilize capillary forces and van der Waals
interactions to coordinate movement and bind microscopic components in
solvent~\cite{Pandey2011, Zhang2017, Siontorou2017}. Additionally,
hydrodynamic interactions are central to fabricate complex,
three-dimensional microstructures~\cite{Dasgupta2017, Leong2007,
Reynolds2019, Cho2010}. Hydrodynamic effects cannot be ignored since
they, like for the rates of biological functions like pore
dynamics~\cite{RYHAM20112929}, set the rate of dissipation. In Specific
Aim 2, we propose novel and efficient high-order numerical methods to
simulate dynamic assembly of particles under the HAP in a viscous
solvent. As described in~\eqref{eq:stokes} and illustrated in
Figure~\ref{fig:flow_map}, the solvent phase is modeled by the Stokes
equations for an incompressible, viscous fluid, and these equations are
coupled to the screened-Laplace equation~\eqref{eq:SL} through viscous
and hydrophobic stress balance.

%\label{subsec:specific_aim_2}
%\begin{wrapfigure}[14]{r}{0.61\textwidth}
%  \centering{\includegraphics[width=0.60\textwidth]{Figures/domain.pdf}}
%    \vspace{-15pt}
%    \caption{\label{fig:flow_map} \footnotesize A collection of 3D
%    amphiphilic particles suspended in a solvent. The solvent velocity
%    satisfies the Stokes equations, and the hydrophobic forces depend on
%    the solvent activity that satisfies the screened Laplace equation.
%    The geometry is updated by solving the mobility problem.} 
%\end{wrapfigure}
The HAP model requires solving exterior problems for the order
parameter~\eqref{eq:SL} and the Stokes equations~\eqref{eq:stokes}. We
will develop methods based on combinations of volume and layer
potentials, and inexact Newton's method. Numerical methods to solve
these integral equations (IEs) have several advantages over their
PDE-based counterpart: by construction, they satisfy far-field and mass
conservation conditions; only $\partial\Omega$ needs to be discretized
for layer potentials; carefully chosen IEs result in a well-conditioned
linear system that is solved with a mesh-independent number of GMRES
iterations; and high-order or spectral accuracy is attainable with
appropriate quadrature methods. The PIs have demonstrated that IEs are a
powerful tool to simulate two-dimensional suspensions of amphiphilic
particles~\cite{Fu2018_SIAM, FuQuRyYo20}. Moreover, elements of these
formulations have been described in three dimensions~\cite{ying_2006,
manasthesis, rac-gre2016}. This proposal will extend the two-dimensional
methods to more general amphiphilic particle models that
satisfy~\eqref{eq:SL}, and all the methods can naturally be extended to
three dimensions. These methods will be motivated by models and analysis
developed in Specific Aims 1 and 3. \S\ref{subsec:AC} describes how the
non-linear Allen-Cahn equation~\eqref{eq:SL} will be iteratively solved
using an IE formulation. Then, \S\ref{subsec:NumericalIssues} describes
how we will combine preconditioning strategies with fast summation
methods to accelerate the time to solution, and
\S\ref{subsec:timeStepping} describes adaptive time stepping methods.

@@

%\begin{wrapfigure}[27]{l}{1.8in}
%  \vspace{-4pt}
%  \centering
%  \input{figures/doublewell.tikz} 
%  \caption{\label{fig:CA} (a) The double well potential $u_0 =
%  0.5$, $u_1 = 2$, $a = 1$, $b = 0.4$. (b) The solution of the
%  Allen-Cahn equation~\eqref{eq:SL}. (c) The convergence of the exact
%  and inexact Newton's methods.}
%\end{wrapfigure}
Having demonstrated that the inexact Newton method converges in one
dimension, it will be coupled with IE formulations to 
solve~\eqref{eq:SL} in two and three dimensions. We first partition
$\Omega$ into $\Omega_0 = \{\xx \in \Omega \:|\: u(\xx) < (u_0 +
u_1)/2\}$ and $\Omega_1 = \{\xx \in \Omega \:|\: u(\xx) > (u_0 +
u_1)/2\}$, and we let $\Lambda = \{\xx \in \Omega \:|\: u(\xx) = (u_0 +
u_1)/2 \}$ be their common boundary, and we let $G_0$ be the fundamental
solution of $(-\rho^2 \Delta + a)$ and $G_1$ be the fundamental
solution of $(-\rho^2 \Delta + 1)$. Then, the solution
of~\eqref{eq:screenedPoisson} can be written as $u = u^P + u^H$ where
$u^P$ is a volume potential and $u^H$ is a layer potential. The volume
potential is
\begin{align*}
  u^{P}(\xx) = \int_{\Omega_i} G_i(\xx,\yy) J[u^N](\yy)\, d\yy, 
    \quad i = 1,2,
\end{align*}
but it is discontinuous across $\Lambda$ and does not satisfy the
prescribed hydrophobic boundary conditions on the particles $\Gamma_i$.
Assuming the particles are entirely contained in the hydrophobic region,
the boundary conditions will be enforced with the layer potentials
\begin{align*}
  u^{H}(\xx) = \begin{cases}
    \displaystyle\int_{\Lambda} G_0(\xx,\yy) \eta_0(\yy) \, ds_\yy + 
    \displaystyle\int_{\Gamma} \pderiv{}{\nn_\yy} G_0(\xx,\yy)
      \sigma(\yy) \,ds_\yy, \quad & \xx \in \Omega_0, \\
    \displaystyle\int_{\Lambda} G_1(\xx,\yy) \eta_1(\yy) \, ds_\yy,
     & \xx \in \Omega_1.
  \end{cases}
\end{align*}
The density functions $\sigma$, $\eta_0$, and $\eta_1$ are determined so
that $u = u^P + u^H$ satisfies the hydrophobic boundary condition on
$\Gamma = \cup_i \Gamma_i$, and that it is continuous and differentiable
across $\Lambda$. This approach of using IEs to solve elliptic PDEs with
piecewise constant coefficients has successfully been used in other
application such as scattering~\cite{hyu-bar2014, che-cho-cai2018}.
